Generalized Plackett-Luce Likelihoods

### 2022-09-28 inst/jss_4382.tex
further work on issue #157
issue #153

\textcolor{blue}{Response}

@@

\textcolor{blue}{Response: Function {\tt maxp()} can return many
  additional objects in addition to the evaluate; standard idiom is to
  pass argument {\tt give = TRUE}.  In particular, a list with
  elements
  \begin{itemize}
  \item {\tt \$hessian} the Hessian matrix: passing argument {\tt give
    = TRUE, hessian = TRUE} returns the Hessian matrix.  However, this
    is restricted to {\tt hyper2} objects and I would add that {\tt
      hyper3} objects are much, much, harder to differentiate than
    {\tt hyper2} objects (let alone second derivatives).
    \item {\tt \$likes} likelihoods
    \item {\tt \$par} the evaluate
  \end{itemize}
  The package includes summary methods and print method for summary
  objects.\\ Some of the suggestions are good for regression-type
  analysis, but do not really carry over to the current
  likelihood-based approach.  The {\tt hyper2} package is predicated
  on straightforward idiom to create and manipulate likelihood
  functions; statistical tests are (to me) just one application of
  likelihood functions.}

@@

\textcolor{blue}{Response: the pattern has been removed as suggested,
  all functions exported are exported explicitly.  However, it is not
  clear who ``the user" is.  Apart from myself, it seems that most
  users of the package are my graduate students who are mostly working
  on applying hyper2 constructions to their own pet datasets; this
  group of users needs to see similar work that they can modify for
  their own projects.  The package was reviewed by Turner in 2020 but
  it is not clear whether she was a user in the sense of the
  comment. In any event, the output of {\tt help(package = "hyper2")}
  is now restricted only to functions that might be of interest to
  users.  }

@@

\textcolor{blue}{Response:}

@@

o Please make sure that all software packages are \cite{}'d properly.

@@

\textcolor{blue}{We believe that this referee's insightful and
  constructive comments have led to a much improved submission.}

>>>>

\textcolor{blue}{Response: new sections 1.1 and 1.2 added discussing
  the new functionality in the package}

@@

\textcolor{blue}{Response: Function {\tt maxp()} can return many
  additional objects in addition to the evaluate; standard idiom is to
  pass argument {\tt give = TRUE}.  In particular, a it returns a list with
  elements including
  \begin{itemize}
  \item {\tt \$hessian} the Hessian matrix: passing argument {\tt give
    = TRUE, hessian = TRUE} returns the Hessian matrix.  However, this
    is restricted to {\tt hyper2} objects and I would add that {\tt
      hyper3} objects are much, much, harder to differentiate than
    {\tt hyper2} objects (let alone second derivatives).
    \item {\tt \$likes} likelihoods
    \item {\tt \$par} the evaluate
  \end{itemize}
  The package includes summary methods and print method for summary
  objects.\\ Some of the suggestions are good for regression-type
  analysis, but do not really carry over to the current
  likelihood-based approach.  The {\tt hyper2} package is predicated
  on straightforward idiom to create and manipulate likelihood
  functions; statistical tests are (to me) just one application of
  likelihood functions.}

@@

\textcolor{blue}{Response: the pattern has been removed as suggested,
  all functions exported are exported explicitly.  However, it is not
  clear who ``the user" is.  The package is still under active
  development.  Apart from myself, it seems that most users of the
  package are my graduate students who are mostly working on applying
  hyper2 constructions to their own pet datasets; this group of users
  needs to see similar work that they can modify for their own
  projects.  The package was reviewed by Turner in 2020 but it is not
  clear whether she was a user in the sense of the comment. In any
  event, the output of {\tt help(package = "hyper2")} is now
  restricted only to functions that might be of interest to users.  }

@@

@@

o Please make sure that all software packages are cited properly.

@@

\textcolor{blue}{Summary: this referee's insightful and
  constructive comments have led to a much improved submission.}

### 2022-09-28 inst/jss_hyper3.Rnw
issue #153

<<>>=
loglik(c(a=0.01, b=0.01, c=0.98),LL)
loglik(c(a=0.90, b=0.05, c=0.05),LL)
@

@@

<<>>=
pchisq(-2 * S_delta, df=1, lower.tail=FALSE)
@

@@

\begin{figure}[htbp]
  \begin{center}
<<figaba>>=
a <- seq(from=0, by=0.005, to=1)
S <- function(a){log(a * (1 - a) / ((1 + a) * (3 - 2 * sqrt(2))))}
plot(a, S(a), type='b')
abline(h=c(0, -2))
abline(v=c(0.02438102, 0.9524271), col='red')
abline(v=sqrt(2) - 1)
@
\caption{A\label{ABA_likelihood} support function for $p_a$ with observation $a\succ b\succ a$}
\end{center}
\end{figure}

@@

p <- function(f, ...){points(a, f(a) / max(f(a)), ...)}
plot(0:1, 0:1, xlab="p(a)", ylab="Likelihood", type="n")
p(f_aab, type="l", col="black")
p(f_aba, type="l", col="red")
p(f_baa, type="l", col="blue")
legend(x=0.85, y=0.7, col=c("black", "red", "blue"), lty=1, pch=NA, legend=c("AAB", "ABA", "BAA"))
abline(h=exp(-2), lty=2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
<<plotwilcoxlike,echo=FALSE>>=
a <- seq(from=0.02, to=0.8, len=40)
L <- sapply(a, function(p){loglik(p, Hxy)})
plot(a, L - max(L),type='b')
abline(h=c(0, -2))
abline(v=c(0.24))
abline(v=c(0.5), lty=2)
@
\caption{A\label{wilcox_likelihood} support for the Bradley-Terry
strength $p_a$ of permeability at term.  The evaluate of 0.24 is
shown; and the two-units-of support credible interval, which does not
exclude $H_0\colon p_a=0.5$ (dotted line), is also shown}
\end{center}
\end{figure}

\section{Multiple competitors: javelin}

@@

<<label=converttosupp3, cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table,
       decreasing=TRUE, give.supp=FALSE)
options(width=60)
javelin_vector
@

@@

<<dothething2,cache=TRUE>>=
javelin <- ordervec2supp3(v=names(javelin_vector)[!is.na(javelin_vector)])
@

@@

<<setdigits,echo=FALSE>>=
options(digits=3)
@

\begin{figure}[htbp]
  \begin{center}
<<testthejav, cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj, pch=16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths}
\end{center}
\end{figure}

@@

<<havealook,echo=FALSE,cache=TRUE>>=
f <- function(s){
  jj <- specificp.test(javelin, "Vesely", s, n=2)
  p <- jj$null_estimate
 return(c(log(p[6] / p[7]), jj$null_support))
}
Ves <- seq(from=0.0199, to=0.33, len=16)
M <- sapply(Ves, f)
M[2,] <- M[2,] - max(M[2,])
rownames(M) <- c("logcontrast", "support")
@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M), type="b")
abline(h=c(0, -2))
abline(v=0, lty=2)
abline(v=log(0.32062833 / 0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected}
\end{center}
\end{figure}

\section{Formula 1 motor racing: the constructors' championship}\label{f1}

@@

<<maxpconstructor2021, cache=TRUE>>=
const2020 <- ordertable2supp3(constructor_2020_table)
const2021 <- ordertable2supp3(constructor_2021_table)
options(digits=4)
maxp(const2020, n=1)
maxp(const2021, n=1)
@

@@

<<usesamep,cache=TRUE>>=
options(digits=4)
samep.test(H, c("Merc2020", "Merc2021"))
@

>>>>

<<>>=
loglik(c(a = 0.01, b = 0.01, c = 0.98), LL)
loglik(c(a = 0.90, b = 0.05, c = 0.05), LL)
@

@@

<<>>=
pchisq(-2 * S_delta, df = 1, lower.tail = FALSE)
@

@@

\begin{figure}[htbp]
  \begin{center}
<<figaba>>=
a <- seq(from = 0, by = 0.005, to = 1)
S <- function(a){log(a * (1 - a) / ((1 + a) * (3 - 2 * sqrt(2))))}
plot(a, S(a), type = 'b')
abline(h = c(0, -2))
abline(v = c(0.02438102, 0.9524271), col = 'red')
abline(v = sqrt(2) - 1)
@
\caption{A\label{ABA_likelihood} support function for $p_a$ with observation $a\succ b\succ a$.}
\end{center}
\end{figure}

@@

p <- function(f, ...){points(a, f(a) / max(f(a)), ...)}
plot(0:1, 0:1, xlab = "p(a)", ylab = "Likelihood", type = "n")
p(f_aab, type = "l", col = "black")
p(f_aba, type = "l", col = "red")
p(f_baa, type = "l", col = "blue")
legend(x=0.85, y=0.7, col = c("black", "red", "blue"), lty=1, pch=NA,
       legend=c("AAB", "ABA", "BAA"))
abline(h = exp(-2), lty = 2)
@
\caption{Likelihood \label{ABAetc_likelihood} functions for
observations $a\succ a\succ b$, $a\succ b\succ a$, $b\succ a\succ a$.}
\end{center}
\end{figure}

@@

\begin{figure}[htbp]
  \begin{center}
<<plotwilcoxlike,echo=FALSE>>=
a <- seq(from = 0.02, to = 0.8, len = 40)
L <- sapply(a, function(p){loglik(p, Hxy)})
plot(a, L - max(L), type = 'b')
abline(h = c(0, -2))
abline(v = c(0.24))
abline(v=c(0.5), lty=2)
@
\caption{A\label{wilcox_likelihood} support function for the Bradley-Terry
strength $p_a$ of permeability at term.  The evaluate of 0.24 is
shown; and the two-units-of support credible interval, which does not
exclude $H_0\colon p_a=0.5$ (dotted line), is also shown.}
\end{center}
\end{figure}

\section{Multiple competitors: Javelin}

@@

<<label=converttosupp3, cache=TRUE>>=
javelin_vector <- attemptstable2supp3(javelin_table,
       decreasing = TRUE, give.supp = FALSE)
options(width = 60)
javelin_vector
@

@@

<<dothething2,cache=TRUE>>=
javelin <- ordervec2supp3(v = names(javelin_vector)[!is.na(javelin_vector)])
@

@@

<<setdigits,echo=FALSE>>=
options(digits = 3)
@

\begin{figure}[htbp]
  \begin{center}
<<testthejav, cache=TRUE>>=
(mj <- maxp(javelin))
dotchart(mj, pch = 16)
@
\caption{Maximum likelihood estimate for javelin throwers' Bradley-Terry strengths.}
\end{center}
\end{figure}

@@

<<havealook,echo=FALSE,cache=TRUE>>=
f <- function(s){
  jj <- specificp.test(javelin, "Vesely", s, n = 2)
  p <- jj$null_estimate
 return(c(log(p[6] / p[7]), jj$null_support))
}
Ves <- seq(from = 0.0199, to = 0.33, len = 16)
M <- sapply(Ves, f)
M[2,] <- M[2,] - max(M[2,])
rownames(M) <- c("logcontrast", "support")
@

\begin{figure}[htbp]
\begin{center}
<<plottheloglikcont,echo=FALSE>>=
colnames(M) <-  as.character(Ves)
plot(t(M), type = "b")
abline(h = c(0, -2))
abline(v = 0, lty = 2)
abline(v = log(0.32062833 / 0.11402735)) # these from mp
@
\caption{Profile \label{profliklogcont} likelihood for log-contrast
$\log\left(p_{\mbox{\tiny Vad}}/ p_{\mbox{\tiny Ves}}\right)$.  Null
of $p_{\mbox{\tiny Vad}}=p_{\mbox{\tiny Ves}}$ indicated with a dotted
line, and two-units-of-support limit indicated with horizontal lines;
thus the null is not rejected.}
\end{center}
\end{figure}

\section{Formula 1 motor racing: the Constructors' Championship}\label{f1}

@@

<<maxpconstructor2021, cache=TRUE>>=
const2020 <- ordertable2supp3(constructor_2020_table)
const2021 <- ordertable2supp3(constructor_2021_table)
options(digits = 4)
maxp(const2020, n = 1)
maxp(const2021, n = 1)
@

@@

<<usesamep,cache=TRUE>>=
options(digits = 4)
samep.test(H, c("Merc2020", "Merc2021"))
@
